Space-time plots

### 2017-05-06 Rmd/space-time.Rmd
Various updates towards revision, including new versions of figure 3

# Space versus time analysis
```{r, message = FALSE, warning=FALSE, results='hold'}
library(ecoscales)
library(readxl)
# library(stringdist)
library(maptools)
library(RColorBrewer)

p_root <- set_base_path("ecoscales")
p_dat <- full_path(p_root, "external/data/result")
p_calib <- full_path(p_root, "external/data/calibration")
# dat <- as.data.table(read_excel(full_path(p_dat, "merged.xlsx")))
dat <- as.data.table(read_excel(full_path(p_dat, "mergedfffff.xlsx")))
# dato <- as.data.table(read_excel(full_path(p_dat, "archive/mergedfff.xlsx")))
cal <- data.table(read_excel(fp(p_calib, "merged_calibrationfffff.xlsx")))
full <- data.table(read_excel(fp(p_dat, "full_no_treuer_toss.xlsx")))
# full <- data.table(read_excel(fp(p_dat, "full_less_Treuer.xlsx")))

@@

# combine full and calibration
# calr[, samp_dur := NA]
calr[, samp_dur := NA]
calr[, "feature" := NULL] #c("DOI", "feature") := NULL]
calr <- merge(calr, calr_type, by = "DOI")
setcolorder(calr, names(dat2))
datf <- rbind(cbind("dsrc" = 1, dat2), cbind("dsrc" = 2, calr))
doisup <- c(unique(cal$DOI_data_source), unique(dat$DOI_data_source))
doisup <- doisup[!is.na(doisup) & doisup != "NA" & doisup != "-"]
doisup <- unlist(strsplit(doisup, ";"))
```

@@

### Adjustments/fixes
1. if `samp_dur != act_dur` and `t_btwn_samp == 0`, set `act_dur == samp_dur` (actually, no, in many cases there is a good reason to have to `act_dur != samp_dur` (see below))
2. Prior to scaling the data in log10 space, we need to make one adjustment. Time between samples has many 0 values because many observations are simply one-offs.  Set these to an arbitrarily large value which will represent clear separation on axis from high frequency studies. A function is defined here for this purpose, and will be applied after bootstrapping is done.

```{r, message=FALSE, warning=FALSE, results='hide'}
# apply fixes (mostly to account for observer omissions)
# 1. if samp_dur != act_dur and t_btwn_samp == 0, set 
#    act_dur == samp_dur
names(datf)
datf[t_btwn_samp == 0 & (act_dur != samp_dur), ]  
dat[t_btwn_samp == 0 & (act_dur != samp_dur), c(1, 3, 8:13), with = F]  
# function to set t_btwn_samp to 365 * 100000 to indicate once-off studies 
# when t_btwn_samp == 0.  
to_infin_byond <- function(tbtwn) ifelse(tbtwn == 0, 365 * 10000, tbtwn)

@@

```{r, eval=FALSE}
# p <- c(0.64, 0.84, 0.53, 1.21)  # from calibration.Rmd
# p <- c(0.58, 0.82, 0.99, 1.24)  # from calibration.Rmd
##     plot_res   n_sites   act_ext   eff_ext t_btwn_samp  act_dur   eff_dur
## 1: 0.5769693 0.3821107 0.7196559 0.4049972     1.01134 1.253045 0.5284913
p <- c(0.58, 0.72, 0.41, 1.01, 1.25, 0.53)  # from calibration.Rmd

# Function for bootstrapping with uncertainties
vars <- c("plot_res", "act_ext", "eff_ext", "t_btwn_samp", "act_dur", 
          "eff_dur")

resamp_func <- function(dat, vars, p, iter) {
  bout <- lapply(1:iter, function(x) {  # x <- 1
    if((x / 100) %in% 1:100) print(x)
    if(length(p) < length(var)) stop("p must be length of var", call. = FALSE)

@@

set.seed(1)
bootperturb <- resamp_func(datf, vars = vars, p = p, iter = 1000) # 10% perturb
# bootperturb2 <- resamp_func(datf, p = p, iter = 1000)  # 10% perturb
# all(bootperturb[[200]]$plot_res == bootperturb2[[200]]$plot_res)
# rm(bootperturb2)
```

```{r, echo = FALSE, eval=FALSE}
# check
par(mfrow = c(2, 3))
for(i in sample(1:1000, 6)) {
  plot(datf[, eff_ext], datf[, eff_ext] - bootperturb[[i]][, eff_ext], 
       pch = 20)
}
```

@@

# effective extent
### fix this here to have label for 10^-6 (so that limit isn't hard-coded below)
alab2 <- c(expression(paste(NULL<="0.1 m"^2)), expression("1 m"^2),
           expression("10 m"^2), 
           expression("100 m"^2), expression("1000 m"^2), expression("1 ha"),
           expression("10 ha"), expression("100 ha"), expression("1000 ha"), 
           expression(paste(10^4 , " ha")), expression(paste(10^5 , " ha")), 
           expression(paste(10^6 , " ha")), expression(paste(10^7 , " ha")), 
           expression(paste(10^8 , " ha")), expression(paste(10^9 , " ha")), 
           expression(paste(10^10 , " ha")))

### fix this here to have value for 10^-6 (so that limit isn't hard-coded below)
aaxis2 <- cbind.data.frame(
  c(0.1 / 10000, 1 / 10000, 10 / 10000, 100 / 10000, 1000 / 10000, 1, 10, 10^2,
    10^3, 10^4, 10^5, 10^6, 10^7, 10^8, 10^9, 10^10))
aaxis2 <- data.table(aaxis2)
setnames(aaxis2, names(aaxis2), "area")
aaxis2[, logarea := log10(area)]
```

@@

```{r, eval=FALSE}
# length(which(datf$study_type %in% c("remote", "other"))) / nrow(datf)
datf[t_btwn_samp == 0, .N] / datf[, .N]  # 38% of samples are once-offs
datf[act_dur == samp_dur, .N] / nrow(datf)
datf[(act_dur == samp_dur) & t_btwn_samp != 0]
datf[(act_dur != samp_dur) & t_btwn_samp == 0]

@@

# set bounds for histograms
# first set variable limits
vlr <- range(aaxis1$logres)  #for resolution
### fix in scaling section (so that min extent limit isn't hard-coded)
vle <- c(-6, range(aaxis2$logarea)[2])  ### fix

@@

# histograms from bootstrap
### fix in scaling section (so that min extent limit isn't hard-coded)
brksl <- list(aaxis1$logres, c(-6, aaxis2$logarea), taxis1$logdays,
              taxis2$logdays)

vs <- c("res", "ext", "int", "dur")
hdat_bs <- lapply(1:length(bootperturb), function(x) {  # x <- 1
  print(x)
  DT <- bootperturb[[x]]
  hdb <- cbind("res" = log10(DT$plot_res), "ext" = log10(DT$eff_ext), 
               "int" = log10(to_infin_byond(DT$t_btwn_samp)), 
               "dur" = log10(DT$eff_dur))
  hdb <- data.table(hdb)
  hdb[res < vlr[1], res := vlr[1]]  # set min res to 0.01 cm"^2 
  hdb[res > vlr[2], res := vlr[2]]  # set max res to 10000 ha 
  # hdb[, lapply(.SD, range, na.rm = TRUE)]
  ### fix in scaling section (so that this limit isn't hard-coded)
  hdb[ext < vle[1], ext := vle[1]]  # set minimum extent -6 (0.01 m2)
  hdb[ext > vle[2], ext := vle[2]]  # set maximum extent 10 (10^10 ha)
  hists <- lapply(1:length(vs), function(y) {  # y <- 2
    #print(y)
    hdbv <- hdb[[vs[y]]]
    hbrks <- brksl[[y]]
    hdbv[hdbv < hbrks[1]] <- hbrks[1]
    hdbv[hdbv > hbrks[length(hbrks)]] <- hbrks[length(hbrks)]
    h <- hist(hdbv, breaks = hbrks, plot = FALSE)
    h$density <- h$counts / sum(h$counts) * 100
    h
  })
})

hdat_bsdt <- lapply(1:length(vs), function(x) { # x <- 1
  mat <- do.call(rbind, lapply(1:length(hdat_bs), function(y) {
    hdat_bs[[y]][[x]]$density
  }))
  DT <- data.table(mat)
})

@@

## Ordinary histograms
mga <- -0.34
sfigl <- -1.2
cxa <- 0.8
cxl <- 0.9
reds <- brewer.pal(9, name = "Reds")
blues <- brewer.pal(9, name = "Blues")
plot(1:2, pch = 20, col = blues[c(4, 6)])
yl <- c(25, 40)
# pdf("paper/figures/hists.pdf", width = 7, height = 2.5)
pdf("paper/figures/hists.pdf", width = 5, height = 5)
par(mfrow = c(2, 2), mar = c(4.5, 1, 1, 1), oma = c(2, 3, 0, 0))
# resolution
h <- hist(hdat$res, breaks = aaxis1$logres, plot = FALSE)
h$density <- h$counts / sum(h$counts) * 100
plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
     tcl = -0.2, cex.axis = cxa, xlab = "Spatial resolution", 
     ylim = c(0, yl[1]), main = "", col = reds[4], cex.lab = cxl)
axis(1, at = aaxis1$logres, labels = alab1, las = 2, cex.axis = cxa, 
     tcl = -0.2, mgp = c(2, 0.25, mga))
mtext("Percent of observations", side = 2, line = 1.5, cex = 0.8)
mtext("A", side = 3, line = sfigl, cex = 0.8, adj = 0.05)

# extent
### fix in scaling section (so that this limit isn't hard-coded)
h <- hist(hdat$ext, breaks = c(-6, aaxis2$logarea), plot = FALSE)
# h <- hist(hdat$aext, breaks = c(-6, aaxis2$logarea), plot = FALSE)
h$density <- h$counts / sum(h$counts) * 100
plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
     tcl = -0.2, cex.axis = cxa, xlab = "Spatial extent", ylim = c(0, yl[1]), 
     main = "", col = reds[6], cex.lab = cxl)
### fix in scaling section (so that this limit isn't hard-coded)
axis(1, at = c(-6, aaxis2$logarea),  
     labels = c(expression(paste(NULL<="0.01m"^2)), expression(paste(0.1, m^2)),                 alab2[-1]), las = 2, cex.axis = cxa,
     tcl = -0.2, mgp = c(2, 0.25, mga))
mtext("B", side = 3, line = sfigl, cex = 0.8, adj = 0.05)

@@

# duration
# tax <- c(taxis2$logdays[-c(10:11)], mean(taxis2$logdays[c(10:11)]))
h <- hist(hdat$dur, breaks = taxis2$logdays, plot = FALSE)
h$density <- h$counts / sum(h$counts) * 100
plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
     tcl = -0.2, cex.axis = cxa, xlab = "Temporal duration", ylim = c(0, yl[2]), 
     main = "", col = blues[6], cex.lab = cxl)
axis(1, at = taxis2$logdays, labels = tlab2, las = 2, cex.axis = cxa, 
     tcl = -0.2, mgp = c(2, 0.25, mga))
mtext("D", side = 3, line = sfigl, cex = 0.8, adj = 0.05)

dev.off()

### Bootstrapped histograms (now main figure)
qtf <- function(x) c(mean(x), quantile(x, probs = c(0.025, 0.975))) # quant f
xlabs <- c("Spatial resolution", "Spatial extent", "Sampling interval", 
           "Temporal duration")
# axes <- list("aax1" = aaxis1$logres, "aax2" = c(-6, aaxis2$logarea), 
#              "tax1" = taxis1$logdays, "tax2" = taxis2$logdays)
### fix in scaling section (so that this limit isn't hard-coded)
axes2 <- list("aax1" = aaxis1$logres, "aax2" = c(-6, aaxis2$logarea), 
              "tax1" = c(taxis2$logdays[-c(10:11)],
                         mean(taxis2$logdays[c(10:11)])), 
              "tax2" = taxis2$logdays)
### fix in scaling section (so that this limit isn't hard-coded)
axlabs <- list(alab1, c(expression(paste(NULL<="0.01m"^2)),
                        expression(paste(0.1, m^2)), alab2[-1]),
               tlab1[-10], tlab2)
mga <- -0.34
sfigl <- -1.2
cxa <- 0.8
cxl <- 0.9
reds <- brewer.pal(9, name = "Reds")
blues <- brewer.pal(9, name = "Blues")
# plot(1:2, pch = 20, col = blues[c(4, 6)])
yl <- 35
cols <- c(reds[c(4, 6)], blues[c(4, 6)])

@@

# stats
# bootp_dt
# axisl <- list(aaxis1$logres, c(-6, aaxis2$logarea), taxis1$logdays,
#               taxis2$logdays)
# hdatstat <- lapply(1:4, function(x) {
#   h <- hist(hdat[[x]], breaks = brksl[[x]], plot = FALSE)
#   h$density <- h$counts / sum(h$counts) * 100
#   h
# })

# res
# hdatst <- hdatstat[[1]]$density
hdatst <- hdatstat_bt$res$mu
l <- length(hdatst)
sum(hdatst[4:6])  # 53 % between 10cm^2 and 1 m^2
sum(hdatst[1:6])  #  67% < 1m^2
sum(hdatst[7:10]) # 24 % between 1 m^2 and 1 ha
sum(hdatst[11:(l)]) # 9 % 1-10000 ha
# dat[plot_res >= 10000^2]

@@

##############
# Ended here 6/4/2017 
##############
# extent
# hdatst <- hdatstat[[2]]$density
hdatst <- hdatstat_bt$ext$mu
l <- length(hdatst)
sum(hdatst[1:2])  # 36 % < 1 m^2
sum(hdatst[1:6])  # 82% < 1 ha
sum(hdatst[1:7])  # 86% < 10 ha
sum(hdatst[1:8])  # 90 % < 100 ha
sum(hdatst[1:9])  # 92 % < 1000 ha
sum(hdatst[8:9])  # 6 % 10-1000 ha
sum(hdatst[8:l])  # 14% > 10 ha
sum(hdatst[10:l]) # 8.2 % covered an area > 1000 ha
sum(hdatst[12:l]) # 5.1 % covered an area > 100,000 ha
sum(hdatst[13:l]) # 3.9 % covered an area > 1,000,000 ha

@@

# datf[, plot(log10(n_sites), log10(plot_res))]
# dat[like(study_type, "field"), plot(log10(n_sites), log10(plot_res))]
datf[like(study_type, "field"), plot(log10(plot_res), log10(act_ext), 
                                     ylim = c(-10, 10))]

datf[like(study_type, "remote"), 
     points(log10(plot_res), log10(act_ext), pch = 20, col = "red")]
datf[like(study_type, "automated"), 
     points(log10(plot_res), log10(act_ext), pch = 20, col = "blue")]
lines(aaxis1$logres, aaxis2$logarea[-16] )
datf[, .N / nrow(datf) * 100, by = st]

@@

# types of observations
# unique(datf$study_type)
datf[like(study_type, "field"), .N] / nrow(datf)  # 80% field studies
datf[like(study_type, "auto"), .N] / nrow(datf) # 12 % automated
datf[like(study_type, "auto") & log10(t_btwn_samp) < -1.380211, .N] /   length(which(hdat[[3]] < -1.380211))  # 75%

datf[like(study_type, "remote|other"), .N] / nrow(datf) # 6 % automated
datf[like(study_type, "paleo"), .N] / nrow(datf) * 100 # <1 % paleo

@@

# choose which dataset to use
# kdat <- datf  # ordinary
kdat <- bootp_dt  # bootstrap
resdat3 <- data.table(x = log10(to_infin_byond(kdat$t_btwn_samp)), 
                      y = log10(kdat$plot_res))
resdat3 <- resdat3[round(x, 8) != 6.56229286, ]  # remove one-time observations
resdat3$n <- rep(1, nrow(resdat3))
rx <- range(resdat3$x)
rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
rx[1] <- -5
#rx[2] <- 6.56
rx[2] <- 5.57
resdat3[x < rx[1], x := rx[1]]
resdat3[x > rx[2], x := rx[2]]
ry <- range(aaxis1$logres)
ry[1] <- -5  # set limits to 0.01 cm^2
resdat3[y < ry[1], y := ry[1]]
resdat3[y > ry[2], y := ry[2]]
coordinates(resdat3) <- ~x + y

# extent
extdat3 <- data.table(x = log10(kdat$act_dur), 
                      y = log10(kdat$act_ext))
extdat3$n <- rep(1, nrow(extdat3))
ex <- range(taxis2$logdays)
ex <- c(floor(ex[1]), 5.56)#ceiling(ex[2]))
extdat3[x < ex[1], x := ex[1]]
extdat3[x > ex[2], x := ex[2]]
ey <- range(aaxis2$logarea)
extdat3[y < ey[1], y := ey[1]]
extdat3[y > ey[2], y := ey[2]]
coordinates(extdat3) <- ~x + y

# temporal framing
tempdat <- data.table(x = log10(to_infin_byond(kdat$t_btwn_samp)), 
                      y = log10(kdat$act_dur))
tempdat <- tempdat[round(x, 8) != 6.56229286, ]  # remove one-time observations
tempdat$n <- rep(1, nrow(tempdat))
tx <- range(tempdat$x)
tx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
#tx[2] <- 6.56
tx[2] <- 5.57
tempdat[x < rx[1], x := rx[1]]
tempdat[x > rx[2], x := rx[2]]
ty <- range(taxis2$logdays)
ty <- c(floor(ty[1]), 5.56)#ceiling(ex[2]))
tempdat[y < ty[1], y := ty[1]]
tempdat[y > ty[2], y := ty[2]]
coordinates(tempdat) <- ~x + y

>>>>

# Space versus time analysis
```{r, message = FALSE, warning=FALSE, results='hold'}
library(ecoscales)
library(readxl)
# library(stringdist)
library(maptools)
library(RColorBrewer)
library(viridis)

p_root <- set_base_path("ecoscales")
p_dat <- full_path(p_root, "external/data/result")
p_calib <- full_path(p_root, "external/data/calibration")
# dat <- as.data.table(read_excel(full_path(p_dat, "merged.xlsx")))
dat <- as.data.table(read_excel(full_path(p_dat, "mergedffffff.xlsx")))
# dato <- as.data.table(read_excel(full_path(p_dat, "archive/mergedfff.xlsx")))
cal <- data.table(read_excel(fp(p_calib, "merged_calibrationffffff.xlsx")))
full <- data.table(read_excel(fp(p_dat, "full_no_treuer_toss.xlsx")))
# full <- data.table(read_excel(fp(p_dat, "full_less_Treuer.xlsx")))

@@

# combine full and calibration
# calr[, samp_dur := NA]
calr[, samp_dur := NA]
calr[, "feature" := NULL] #c("DOI", "feature") := NULL]
calr <- merge(calr, calr_type, by = "DOI")
setcolorder(calr, names(dat2))
datf <- rbind(cbind("dsrc" = 1, dat2), cbind("dsrc" = 2, calr))
datf[act_dur > eff_dur]
doisup <- c(unique(cal$DOI_data_source), unique(dat$DOI_data_source))
doisup <- doisup[!is.na(doisup) & doisup != "NA" & doisup != "-"]
doisup <- unlist(strsplit(doisup, ";"))
```

@@

### Adjustments/fixes
1. One could argue that if `samp_dur != act_dur` and `t_btwn_samp == 0`, set `act_dur == samp_dur`, but in a small number of cases there is a good reason to have to `act_dur != samp_dur` (e.g. DOI 10.1890/08-0611.1 when farmers were asked about their seasonal management practices--time period of interview was clearly != to information being collected))
2. Prior to scaling the data in log10 space, we need to make one adjustment. Time between samples has many 0 values because many observations are simply one-offs.  Set these to an arbitrarily large value which will represent clear separation on axis from high frequency studies. A function is defined here for this purpose, and will be applied after bootstrapping is done.

```{r, message=FALSE, warning=FALSE, results='hide'}
# apply fixes (mostly to account for observer omissions)
# 1. if samp_dur != act_dur and t_btwn_samp == 0, set 
#    act_dur == samp_dur
names(datf)
datf[t_btwn_samp == 0 & (act_dur != samp_dur), ]  
dat[t_btwn_samp == 0 & (act_dur != samp_dur), c(1, 3:4, 9:15), with = FALSE]  
# function to set t_btwn_samp to 365 * 100000 to indicate once-off studies 
# when t_btwn_samp == 0.  
dat[act_dur < 1 & t_btwn_samp > 1000, ]
dat[act_dur > 1 & t_btwn_samp == 0, c(1, 3:4, 9:15), with = FALSE]
```

@@

```{r, eval=FALSE}
# p <- c(0.64, 0.84, 0.53, 1.21)  # from calibration.Rmd
# p <- c(0.58, 0.82, 0.99, 1.24)  # from calibration.Rmd
##     plot_res   n_sites   act_ext   eff_ext t_btwn_samp  act_dur   eff_dur
## 1: 0.5769693 0.3821107 0.7196559 0.4049972     1.01134 1.253045 0.5284913
p <- c(0.58, 0.75, 0.50, 1.01, 1.23, 0.59)  # from calibration.Rmd

# Functions for bootstrapping with uncertainties
# sets unreplicated observations to arbitrarily large number
to_infin_byond <- function(tbtwn) ifelse(tbtwn == 0, 365 * 10000, tbtwn)

# bootstrapping function
resamp_func <- function(dat, vars, p, iter) {
  bout <- lapply(1:iter, function(x) {  # x <- 1
    if((x / 100) %in% 1:100) print(x)
    if(length(p) < length(var)) stop("p must be length of var", call. = FALSE)

@@

# Run bootstrap
vars <- c("plot_res", "act_ext", "eff_ext", "t_btwn_samp", "act_dur", "eff_dur")
set.seed(1)
bootperturb <- resamp_func(datf, vars = vars, p = p, iter = 1000) # 10% perturb
# bootperturb2 <- resamp_func(datf, p = p, iter = 1000)  # 10% perturb
# all(bootperturb[[200]]$plot_res == bootperturb2[[200]]$plot_res)
# rm(bootperturb2)
```

```{r, echo = FALSE, eval=FALSE}
# check
par(mfrow = c(2, 3))
for(i in sample(1:1000, 6)) {
  plot(datf[, eff_ext], datf[, eff_ext] - bootperturb[[i]][, eff_ext], 
       pch = 20)
}
dev.off()
```

@@

# effective extent
alab2 <- c(expression(paste(NULL<="0.01 m"^2)), expression(paste("0.1 m"^2)), 
           expression("1 m"^2), expression("10 m"^2), 
           expression("100 m"^2), expression("1000 m"^2), expression("1 ha"),
           expression("10 ha"), expression("100 ha"), expression("1000 ha"), 
           expression(paste(10^4 , " ha")), expression(paste(10^5 , " ha")), 
           expression(paste(10^6 , " ha")), expression(paste(10^7 , " ha")), 
           expression(paste(10^8 , " ha")), expression(paste(10^9 , " ha")), 
           expression(paste(10^10 , " ha")))

aaxis2 <- cbind.data.frame(
  c(0.01 / 10000, 0.1 / 10000, 1 / 10000, 10 / 10000, 100 / 10000, 1000 / 10000,
    1, 10, 10^2, 10^3, 10^4, 10^5, 10^6, 10^7, 10^8, 10^9, 10^10))
aaxis2 <- data.table(aaxis2)
setnames(aaxis2, names(aaxis2), "area")
aaxis2[, logarea := log10(area)]
```

@@

Histograms for each of the four main dimensions.
```{r, eval=FALSE}
# length(which(datf$study_type %in% c("remote", "other"))) / nrow(datf)
datf[t_btwn_samp == 0, .N] / datf[, .N]  # 37% of samples are once-offs
datf[act_dur == samp_dur, .N] / nrow(datf)
datf[(act_dur == samp_dur) & t_btwn_samp != 0]
datf[(act_dur != samp_dur) & t_btwn_samp == 0]

@@

# set bounds for histograms
# first set variable limits
vlr <- range(aaxis1$logres)  # for resolution
vle <- range(aaxis2$logarea)  # for extent

@@

# histograms from bootstrap
brksl <- list(aaxis1$logres, aaxis2$logarea, taxis1$logdays, taxis2$logdays)

# Assign bootstrapped values to bins, making necessary adjustments as needed 
vs <- c("res", "ext", "int", "dur")
hdat_bs <- lapply(1:length(bootperturb), function(x) {  # x <- 1
  print(x)
  DT <- bootperturb[[x]]
  hdb <- cbind("res" = log10(DT$plot_res), "ext" = log10(DT$eff_ext), 
               "int" = log10(to_infin_byond(DT$t_btwn_samp)), 
               "dur" = log10(DT$eff_dur))
  hdb <- data.table(hdb)
  hdb[res < vlr[1], res := vlr[1]]  # set min res to 0.01 cm"^2 
  hdb[res > vlr[2], res := vlr[2]]  # set max res to 10000 ha 
  # hdb[, lapply(.SD, range, na.rm = TRUE)]
  hdb[ext < vle[1], ext := vle[1]]  # set minimum extent -6 (0.01 m2)
  hdb[ext > vle[2], ext := vle[2]]  # set maximum extent 10 (10^10 ha)
  hists <- lapply(1:length(vs), function(y) {  # y <- 2
    #print(y)
    hdbv <- hdb[[vs[y]]]
    hbrks <- brksl[[y]]
    hdbv[hdbv < hbrks[1]] <- hbrks[1]
    hdbv[hdbv > hbrks[length(hbrks)]] <- hbrks[length(hbrks)]
    h <- hist(hdbv, breaks = hbrks, plot = FALSE)
    h$density <- h$counts / sum(h$counts) * 100
    h
  })
})

# convert density values for each dimension to data.tables
hdat_bsdt <- lapply(1:length(vs), function(x) { # x <- 1
  mat <- do.call(rbind, lapply(1:length(hdat_bs), function(y) {
    hdat_bs[[y]][[x]]$density
  }))
  DT <- data.table(mat)
})

@@

## Ordinary histograms
mga <- -0.34
sfigl <- -1.2
cxa <- 0.8
cxl <- 0.9
reds <- brewer.pal(9, name = "Reds")
blues <- brewer.pal(9, name = "Blues")
# plot(1:2, pch = 20, col = blues[c(4, 6)])
yl <- c(25, 40)
# pdf("paper/figures/hists.pdf", width = 7, height = 2.5)
pdf("paper/figures/hists.pdf", width = 5, height = 5)
par(mfrow = c(2, 2), mar = c(4.5, 1, 1, 1), oma = c(2, 3, 0, 0))
# resolution
h <- hist(hdat$res, breaks = aaxis1$logres, plot = FALSE)
h$density <- h$counts / sum(h$counts) * 100
plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
     tcl = -0.2, cex.axis = cxa, xlab = "Spatial resolution", 
     ylim = c(0, yl[1]), main = "", col = reds[4], cex.lab = cxl)
axis(1, at = aaxis1$logres, labels = alab1, las = 2, cex.axis = cxa, 
     tcl = -0.2, mgp = c(2, 0.25, mga))
mtext("Percent of observations", side = 2, line = 1.5, cex = 0.8)
mtext("A", side = 3, line = sfigl, cex = 0.8, adj = 0.05)

# extent
h <- hist(hdat$ext, breaks = aaxis2$logarea, plot = FALSE)
# h <- hist(hdat$aext, breaks = c(-6, aaxis2$logarea), plot = FALSE)
h$density <- h$counts / sum(h$counts) * 100
plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
     tcl = -0.2, cex.axis = cxa, xlab = "Spatial extent", ylim = c(0, yl[1]), 
     main = "", col = reds[6], cex.lab = cxl)
axis(1, at = aaxis2$logarea,
     labels = alab2,
     # labels = c(expression(paste(NULL<="0.01m"^2)), expression(paste(0.1, m^2)),                 alab2[-1]), 
     las = 2, cex.axis = cxa,
     tcl = -0.2, mgp = c(2, 0.25, mga))
mtext("B", side = 3, line = sfigl, cex = 0.8, adj = 0.05)

@@

# duration
# tax <- c(taxis2$logdays[-c(10:11)], mean(taxis2$logdays[c(10:11)]))
h <- hist(hdat$dur, breaks = taxis2$logdays, plot = FALSE)
h$density <- h$counts / sum(h$counts) * 100
plot(h, freq = FALSE, ylab = "", las = 2,  xaxt = "n", mgp = c(3.25, 0.25, mga),
     tcl = -0.2, cex.axis = cxa, xlab = "Temporal duration", ylim = c(0, yl[2]),      main = "", col = blues[6], cex.lab = cxl)
axis(1, at = taxis2$logdays, labels = tlab2, las = 2, cex.axis = cxa, 
     tcl = -0.2, mgp = c(2, 0.25, mga))
mtext("D", side = 3, line = sfigl, cex = 0.8, adj = 0.05)
dev.off()

### Bootstrapped histograms (now main figure)
qtf <- function(x) c(mean(x), quantile(x, probs = c(0.025, 0.975))) # quant f
xlabs <- c("Spatial resolution", "Spatial extent", "Sampling interval", 
           "Temporal duration")
# axes <- list("aax1" = aaxis1$logres, "aax2" = c(-6, aaxis2$logarea), 
#              "tax1" = taxis1$logdays, "tax2" = taxis2$logdays)
axes2 <- list("aax1" = aaxis1$logres, "aax2" = aaxis2$logarea, 
              "tax1" = c(taxis2$logdays[-c(10:11)],
                         mean(taxis2$logdays[c(10:11)])), 
              "tax2" = taxis2$logdays)
# axlabs <- list(alab1, c(expression(paste(NULL<="0.01m"^2)),
#                         expression(paste(0.1, m^2)), alab2[-c(1:2)]),
axlabs <- list(alab1, alab2, tlab1[-10], tlab2)
mga <- -0.34
sfigl <- -1.2
cxa <- 0.8
cxl <- 0.9
reds <- brewer.pal(9, name = "Reds")
blues <- brewer.pal(9, name = "Blues")
# plot(1:2, pch = 20, col = blues[c(4, 6)])
yl <- 35
cols <- c(reds[c(4, 6)], blues[c(4, 6)])

@@

# stats
# bootp_dt
# axisl <- list(aaxis1$logres, c(-6, aaxis2$logarea), taxis1$logdays,
#               taxis2$logdays)
# hdatstat <- lapply(1:4, function(x) {
#   h <- hist(hdat[[x]], breaks = brksl[[x]], plot = FALSE)
#   h$density <- h$counts / sum(h$counts) * 100
#   h
# })
```

Statistics associated with Figure 1: percentage of observations falling within different scale ranges, and associated with different observational methods
```{r}
# res
# hdatst <- hdatstat[[1]]$density
hdatst <- hdatstat_bt$res$mu
l <- length(hdatst)
sum(hdatst[4:6])  # 53 % between 10cm^2 and 1 m^2
sum(hdatst[1:6])  #  67% < 1m^2
sum(hdatst[7:10]) # 24 % between 1 m^2 and 1 ha
sum(hdatst[11:(l)]) # 9 % 1-10000 ha
# dat[plot_res >= 10000^2]

@@

# extent
# hdatst <- hdatstat[[2]]$density
hdatst <- hdatstat_bt$ext$mu
l <- length(hdatst)
sum(hdatst[1:2])  # 36 % < 1 m^2
sum(hdatst[1:6])  # 82% < 1 ha
sum(hdatst[1:7])  # 86% < 10 ha
sum(hdatst[1:8])  # 90 % < 100 ha
sum(hdatst[1:9])  # 92 % < 1000 ha
sum(hdatst[8:9])  # 6 % 10-1000 ha
sum(hdatst[8:l])  # 14% > 10 ha
sum(hdatst[10:l]) # 8.2 % covered an area > 1000 ha
sum(hdatst[12:l]) # 5.1 % covered an area > 100,000 ha
sum(hdatst[13:l]) # 3.9 % covered an area > 1,000,000 ha

@@

# datf[, plot(log10(n_sites), log10(plot_res))]
# dat[like(study_type, "field"), plot(log10(n_sites), log10(plot_res))]
datf[like(study_type, "field"), plot(log10(plot_res), log10(act_ext), 
                                     ylim = c(-10, 10))]
datf[like(study_type, "remote"), 
     points(log10(plot_res), log10(act_ext), pch = 20, col = "red")]
datf[like(study_type, "automated"), 
     points(log10(plot_res), log10(act_ext), pch = 20, col = "blue")]
lines(aaxis1$logres, aaxis2$logarea[-c(16:17)] )
datf[, .N / nrow(datf) * 100, by = st]

@@

# types of observations
# unique(datf$study_type)
datf[like(study_type, "field"), .N] / nrow(datf)  # 80% field studies
datf[like(study_type, "auto"), .N] / nrow(datf) # 12 % automated
datf[like(study_type, "auto") & log10(t_btwn_samp) < -1.380211, .N] / 
  length(which(hdat[["int"]] < -1.380211))  # 78 % high frequency obs automated 
datf[like(study_type, "remote|other"), .N] / nrow(datf) # 7 % remote
datf[like(study_type, "paleo"), .N] / nrow(datf) * 100 # <1 % paleo

@@

# choose which dataset to use
# kdat <- datf  # ordinary
kdat <- bootp_dt  # bootstrap
resdat3 <- data.table(x = log10(to_infin_byond(kdat$t_btwn_samp)), 
                      y = log10(kdat$plot_res))
resdat3 <- resdat3[round(x, 8) != 6.56229286]  # remove one-time observations
resdat3[, n := 1]  # set up count column
# resdat3$n <- rep(1, nrow(resdat3))
rx <- range(resdat3$x)
rx <- round(rx / 0.5) * 0.5  # round to nearest 0.5
rx[1] <- -5  # set min interval to less than second
#rx[2] <- 6.56
rx[2] <- 5.57  # set max interval to millenium
resdat3[x < rx[1], x := rx[1]]
resdat3[x > rx[2], x := rx[2]]
ry <- range(aaxis1$logres)
ry[1] <- -5  # set limits to 0.01 cm^2
resdat3[y < ry[1], y := ry[1]]
resdat3[y > ry[2], y := ry[2]]
coordinates(resdat3) <- ~x + y

# extent
extdat3 <- data.table(x = log10(kdat$eff_dur),#log10(kdat$act_dur), 
                      y = log10(kdat$eff_ext))#log10(kdat$act_ext))
extdat3 <- extdat3[!is.na(x) & !is.na(y)] # cut out null values
# extdat3$n <- rep(1, nrow(extdat3))
extdat3[, n := 1]  # set up count column
ex <- range(taxis2$logdays)
ex <- c(floor(ex[1]), 5.56) ### should this be 5.57?  #ceiling(ex[2]))
extdat3[x < ex[1], x := ex[1]]
extdat3[x > ex[2], x := ex[2]]
ey <- range(aaxis2$logarea)
extdat3[y < ey[1], y := ey[1]]
extdat3[y > ey[2], y := ey[2]]
coordinates(extdat3) <- ~x + y
